Forecasting the 2024 U.S. Presidential Election: An Analysis for Battleground States\thanks{Code and data are available at: \url{https://github.com/krishnak30/US_elections}.}

### 2024-11-01 paper/paper.qmd
cleaning files

bibliography: references.bib

---

@@

This paper is organized as follows: In {#sec-data}, we clean the data, explore summary statistics, plot distributions of key variables, and examine relationships between variables. In {#sec-model}, we discuss our forecasting approach, model selection, justification for the chosen model, and the mechanism of deriving poll weights based on pollster reliability, ultimately presenting our predictions. Finally, in {#sec-discussion}, we address the broader implications of our findings, acknowledge limitations, and suggest directions for future work.
\vspace{0.7cm}
# Data {#sec-data}
\vspace{0.7cm}
## Measurement

@@

This methodology outlines an election forecasting plan with a budget of \$100,000. Heavily inspired by TIPP Insights and Emerson College's hybrid approaches, the survey methodology balances cost-efficiency and data accuracy by using different modes of data collection and wide demographic reach.

This forecasting plan will predict the popular vote and Electoral College outcomes, emphasizing key swing states and underrepresented voter groups.

@@

This idealized methodology uses a multi-mode hybrid sampling approach. This allows the use of both probability-based and non-probability-based methods:

-   Probability-based methods:

    -   Live phone interviews (landline and cell): As used by TIPP, live phone interviews is a direct and personal method of collecting responses. Voter lists would be sourced from Aristotle to ensure we contact likely voters. This mode reaches older voters and voters with less internet connection.

    -   Interactive Voice Response (IVR): Similar to Emerson’s approach, IVR for landline users balances costs and ensures demographic representation, particularly among older or less tech-savvy populations. This method captures quick responses and is less expensive than live interviews.

-   Non-probability methods:

    -   MMS-to-web (mobile): Following Emerson’s MMS-to-web strategy, we would send multimedia messages to likely voters’ cell phones, encouraging them to complete an online survey. This method works well for younger, mobile-first voters, helping fill in gaps where traditional methods might miss.

    -   Online panel recruitment via CINT: To mirror TIPP’s online panel approach, a portion of the survey would be conducted via an opt-in panel of pre-screened voters from CINT, capturing tech-savvy respondents. CINT’s platform ensures that different voter groups, including hard-to-reach groups, are represented in the polling.

@@

By using Aristotle’s voter lists and CINT's platform as a sampling frame, the sample consists of likely voters based on registration status and past voting behavior. Quota sampling would be used to match the demographic profile of the U.S. electorate. This ensures that respondents in the sample reflect key demographic variables such as age, gender, race, and education level.

We would oversample key swing states like Pennsylvania, Wisconsin, and Florida. Key swing states are states where there is no overwhelming support for a particular candidate or political party. The outcome in these states often determine the overall result of the election, making them critical in securing a win in the Electoral College. Oversampling these states allows a more accurate forecasting in these battlegrounds, which are critical for predicting the Electoral College outcome.

@@

Following TIPP’s iterative weighting process, the data will be weighted to ensure that the sample aligns with key demographics (race, gender, education) based on census data.

>>>>

bibliography: references.bib
---

@@

This paper is organized as follows: In \ref{sec-data}, we clean the data, explore summary statistics, plot distributions of key variables, and examine relationships between variables. In \ref{sec-model}, we discuss our forecasting approach, model selection, justification for the chosen model, and the mechanism of deriving poll weights based on pollster reliability, ultimately presenting our predictions. Finally, in \ref{sec-discussion}, we address the broader implications of our findings, acknowledge limitations, and suggest directions for future work. \vspace{0.7cm} \# Data {#sec-data} \vspace{0.7cm} \## Measurement

@@

This methodology outlines an election forecasting plan with a budget of \$100,000. This survey is heavily inspired by TIPP Insights and Emerson College Polling (ECP), which are two respected polling institutions. TIPP Insights is known for accurately predicting major U.S. elections, including presidential races, by combining traditional live phone interviews with probability-based sampling [@tipp_accuracy]. Their history of reliable forecasts has made TIPP a trusted name in polling, that specializes in capturing voter sentiment and representing diverse demographics. Similarly, Emerson College is known for its mixed-mode polling strategy, which combines Interactive Voice Response (IVR) with online polling to reach both older, landline users and younger, mobile-first voters [@about_ecp].

By incorporating techniques from both of these institutions, our methodology aims to achieve high accuracy and broad demographic reach, ensuring a balanced, cost-effective approach to election forecasting.

@@

This idealized methodology employs a multi-mode hybrid sampling approach, utilizing both probability-based and non-probability-based methods. Combining both methods provides both accuracy and inclusivity in polling. Overall, this method creates a more complete view of voter preferences [@baker].

Probability-based methods include live phone interviews (landline and cell), as used by TIPP. Live interviews are a direct, personal method of collecting responses, with voter lists sourced from Aristotle [@tipp_methodology]. This method is effective in reaching older voters and voters with less internet connectivity. Interactive Voice Response (IVR) is also used, following Emerson’s strategy for landline users [@about_ecp], to balance costs and ensure representation of less tech-savvy populations.

For non-probability methods, this approach includes MMS-to-web surveys for mobile users, inspired by Emerson’s MMS-to-web strategy [@about_ecp]. Likely voters receive multimedia messages on their cell phones encouraging them to complete an online survey, which helps reach younger, mobile-first voters. Additionally, online panel recruitment is done through CINT. This follows TIPP's online panel strategy, where a part of the survey is delivered to a pre-screened panel [@tipp_methodology].

The combination of these sampling methods and collection methods ensures a more accurate representation of voter groups, including harder-to-reach populations.

@@

By using Aristotle’s voter lists and CINT's platform as a sampling frame, the sample consists of likely voters. Likely voters are determined based on registration status and past voting behavior [@ecp_poll]. Quota sampling would be used to match the demographic profile of the U.S. electorate. Quota sampling is a sampling method that relies on a non-random selection of units based on predetermined proportions [@def_quota]. This ensures that respondents in the sample reflect demographic variables such as age, gender, race, and education level.

We would oversample key swing states like Pennsylvania, Arizona, and Georgia [@swing_state]. Key swing states are states where there is no overwhelming support for a particular candidate or political party [@barker]. The outcome in these states often determine the overall result of the election, making them critical in securing a win in the Electoral College. Oversampling these states allows a more accurate forecasting in these battlegrounds, which are critical for predicting the Electoral College outcome.

@@

The data will be weighted to ensure that the sample aligns with key demographics (race, gender, education) based on census data. Weights are applied to even out under- and over-represented groups, making the data more similar to the electorate [@weight].

### 2024-11-01 paper/paper.tex
cleaning files

\begin{document}
\maketitle
\begin{abstract}
This study employs a polls-of-polls approach to predict support for
Kamala Harris and Donald Trump in the 2024 U.S. Presidential Election,
with a focus on key battleground states to forecast the likely winner of
the electoral college. By aggregating multiple polls and applying a
weighted linear regression model with predictors like pollster
reliability, sample size, state, and recency, we estimate a higher
national support for Trump relative to Harris. Our analysis also reveals
a close competition in battleground states, with Trump holding a narrow
lead in Arizona, Georgia, and North Carolina, while Harris leads in
Michigan, Nevada, Pennsylvania, and Wisconsin. These findings underscore
the value of aggregating polls over relying on individual surveys,
offering a more comprehensive and robust forecast of electoral outcomes.
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, sharp corners, interior hidden, enhanced, borderline west={3pt}{0pt}{shadecolor}, frame hidden, breakable]}{\end{tcolorbox}}\fi

@@

This paper is organized as follows: In \{\#sec-data\}, we clean the
data, explore summary statistics, plot distributions of key variables,
and examine relationships between variables. In \{\#sec-model\}, we
discuss our forecasting approach, model selection, justification for the
chosen model, and the mechanism of deriving poll weights based on
pollster reliability, ultimately presenting our predictions. Finally, in
\{\#sec-discussion\}, we address the broader implications of our
findings, acknowledge limitations, and suggest directions for future
work. \vspace{0.7cm} \# Data \{\#sec-data\} \vspace{0.7cm} \#\#
Measurement

@@

This methodology outlines an election forecasting plan with a budget of
\$100,000. Heavily inspired by TIPP Insights and Emerson College's
hybrid approaches, the survey methodology balances cost-efficiency and
data accuracy by using different modes of data collection and wide
demographic reach.

This forecasting plan will predict the popular vote and Electoral
College outcomes, emphasizing key swing states and underrepresented
voter groups.

@@

This idealized methodology uses a multi-mode hybrid sampling approach.
This allows the use of both probability-based and non-probability-based
methods:

\begin{itemize}
\item
  Probability-based methods:

  \begin{itemize}
  \item
    Live phone interviews (landline and cell): As used by TIPP, live
    phone interviews is a direct and personal method of collecting
    responses. Voter lists would be sourced from Aristotle to ensure we
    contact likely voters. This mode reaches older voters and voters
    with less internet connection.
  \item
    Interactive Voice Response (IVR): Similar to Emerson's approach, IVR
    for landline users balances costs and ensures demographic
    representation, particularly among older or less tech-savvy
    populations. This method captures quick responses and is less
    expensive than live interviews.
  \end{itemize}
\item
  Non-probability methods:

  \begin{itemize}
  \item
    MMS-to-web (mobile): Following Emerson's MMS-to-web strategy, we
    would send multimedia messages to likely voters' cell phones,
    encouraging them to complete an online survey. This method works
    well for younger, mobile-first voters, helping fill in gaps where
    traditional methods might miss.
  \item
    Online panel recruitment via CINT: To mirror TIPP's online panel
    approach, a portion of the survey would be conducted via an opt-in
    panel of pre-screened voters from CINT, capturing tech-savvy
    respondents. CINT's platform ensures that different voter groups,
    including hard-to-reach groups, are represented in the polling.
  \end{itemize}
\end{itemize}

@@

By using Aristotle's voter lists and CINT's platform as a sampling
frame, the sample consists of likely voters based on registration status
and past voting behavior. Quota sampling would be used to match the
demographic profile of the U.S. electorate. This ensures that
respondents in the sample reflect key demographic variables such as age,
gender, race, and education level.

We would oversample key swing states like Pennsylvania, Wisconsin, and
Florida. Key swing states are states where there is no overwhelming
support for a particular candidate or political party. The outcome in
these states often determine the overall result of the election, making
them critical in securing a win in the Electoral College. Oversampling
these states allows a more accurate forecasting in these battlegrounds,
which are critical for predicting the Electoral College outcome.

@@

Following TIPP's iterative weighting process, the data will be weighted
to ensure that the sample aligns with key demographics (race, gender,
education) based on census data.

@@

@@

@@

@@

@@

>>>>

\begin{document}
\maketitle
\begin{abstract}
This study employs a polls-of-polls approach to predict support for
Kamala Harris and Donald Trump in the 2024 U.S. Presidential Election,
with a focus on key battleground states to forecast the likely winner of
the electoral college. By aggregating multiple polls and applying a
weighted linear regression model with predictors like pollster
reliability, sample size, state, and recency, we estimate a higher
national support for Trump relative to Harris. Our analysis also reveals
a close competition in battleground states, with Trump holding a narrow
lead in Arizona, Georgia, and North Carolina, while Harris leads in
Michigan, Nevada, Pennsylvania, and Wisconsin. These findings underscore
the value of aggregating polls over relying on individual surveys,
offering a more comprehensive and robust forecast of electoral outcomes.
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, breakable, boxrule=0pt, sharp corners, interior hidden, frame hidden, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi

@@

This paper is organized as follows: In \ref{sec-data}, we clean the
data, explore summary statistics, plot distributions of key variables,
and examine relationships between variables. In \ref{sec-model}, we
discuss our forecasting approach, model selection, justification for the
chosen model, and the mechanism of deriving poll weights based on
pollster reliability, ultimately presenting our predictions. Finally, in
\ref{sec-discussion}, we address the broader implications of our
findings, acknowledge limitations, and suggest directions for future
work. \vspace{0.7cm} \# Data \{\#sec-data\} \vspace{0.7cm} \#\#
Measurement

@@

This methodology outlines an election forecasting plan with a budget of
\$100,000. This survey is heavily inspired by TIPP Insights and Emerson
College Polling (ECP), which are two respected polling institutions.
TIPP Insights is known for accurately predicting major U.S. elections,
including presidential races, by combining traditional live phone
interviews with probability-based sampling (Rakich 2023). Their history
of reliable forecasts has made TIPP a trusted name in polling, that
specializes in capturing voter sentiment and representing diverse
demographics. Similarly, Emerson College is known for its mixed-mode
polling strategy, which combines Interactive Voice Response (IVR) with
online polling to reach both older, landline users and younger,
mobile-first voters (Emerson College Polling 2024a).

By incorporating techniques from both of these institutions, our
methodology aims to achieve high accuracy and broad demographic reach,
ensuring a balanced, cost-effective approach to election forecasting.

@@

This idealized methodology employs a multi-mode hybrid sampling
approach, utilizing both probability-based and non-probability-based
methods. Combining both methods provides both accuracy and inclusivity
in polling. Overall, this method creates a more complete view of voter
preferences (Baker et al. 2013).

Probability-based methods include live phone interviews (landline and
cell), as used by TIPP. Live interviews are a direct, personal method of
collecting responses, with voter lists sourced from Aristotle (TIPP
Staff 2023). This method is effective in reaching older voters and
voters with less internet connectivity. Interactive Voice Response (IVR)
is also used, following Emerson's strategy for landline users (Emerson
College Polling 2024a), to balance costs and ensure representation of
less tech-savvy populations.

For non-probability methods, this approach includes MMS-to-web surveys
for mobile users, inspired by Emerson's MMS-to-web strategy (Emerson
College Polling 2024a). Likely voters receive multimedia messages on
their cell phones encouraging them to complete an online survey, which
helps reach younger, mobile-first voters. Additionally, online panel
recruitment is done through CINT. This follows TIPP's online panel
strategy, where a part of the survey is delivered to a pre-screened
panel (TIPP Staff 2023).

The combination of these sampling methods and collection methods ensures
a more accurate representation of voter groups, including
harder-to-reach populations.

@@

By using Aristotle's voter lists and CINT's platform as a sampling
frame, the sample consists of likely voters. Likely voters are
determined based on registration status and past voting behavior
(Emerson College Polling 2024b). Quota sampling would be used to match
the demographic profile of the U.S. electorate. Quota sampling is a
sampling method that relies on a non-random selection of units based on
predetermined proportions (Nikolopoulou 2022). This ensures that
respondents in the sample reflect demographic variables such as age,
gender, race, and education level.

We would oversample key swing states like Pennsylvania, Arizona, and
Georgia (FitzGerald 2024). Key swing states are states where there is no
overwhelming support for a particular candidate or political party
(Barket 2020). The outcome in these states often determine the overall
result of the election, making them critical in securing a win in the
Electoral College. Oversampling these states allows a more accurate
forecasting in these battlegrounds, which are critical for predicting
the Electoral College outcome.

@@

The data will be weighted to ensure that the sample aligns with key
demographics (race, gender, education) based on census data. Weights are
applied to even out under- and over-represented groups, making the data
more similar to the electorate (Elliott 2020).

@@

\leavevmode\vadjust pre{\hypertarget{ref-baker}{}}%
Baker, Reg, Michael Brick, Nancy Bates, Mike Battaglia, Mick Couper,
Jill Dever, Krista Gile, and Roger Tourangeau. 2013. \emph{REPORT OF THE
AAPOR TASK FORCE ON NONPROBABILITY SAMPLING}.
\url{https://aapor.org/wp-content/uploads/2022/11/NPS_TF_Report_Final_7_revised_FNL_6_22_13-2.pdf}.

\leavevmode\vadjust pre{\hypertarget{ref-barker}{}}%
Barket, Elaine. 2020. \emph{Recommendation for Key Management: Part 1 --
General}.
\url{https://nvlpubs.nist.gov/nistpubs/SpecialPublications/NIST.SP.800-57pt1r5.pdf}.

@@

\leavevmode\vadjust pre{\hypertarget{ref-swing_state}{}}%
FitzGerald, James. 2024. \emph{The Seven Swing States Set to Decide the
2024 US Election}. \url{https://www.bbc.com/news/articles/c511pyn3xw3o}.

@@

\leavevmode\vadjust pre{\hypertarget{ref-def_quota}{}}%
Nikolopoulou, Kassiani. 2022. \emph{What Is Quota Sampling? \textbar{}
Definition \& Examples}.
\url{https://www.scribbr.com/methodology/quota-sampling/}.

@@

\leavevmode\vadjust pre{\hypertarget{ref-tipp_accuracy}{}}%
Rakich, Nathaniel. 2023. \emph{The Polls Were Historically Accurate in
2022}.
\url{https://fivethirtyeight.com/features/2022-election-polling-accuracy/}.

@@

\leavevmode\vadjust pre{\hypertarget{ref-tipp_methodology}{}}%
TIPP Staff. 2023. \emph{TIPP Indexes Methodology}.
\url{https://tippinsights.com/tipp-indexes-methodology/}.
